---
title: "COVID-19 Analysis"
author: "iffig"
date: '2022-06-13'
output: html_document
---

# Objective 
tbd
# Data Preparation 

## Importing 
```{r libraryImports, echo=FALSE}
library(tidyverse)
library(lubridate)
```

```{r importData, echo=FALSE}
url <- "https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/"
files <- c("time_series_covid19_confirmed_US.csv", 
           "time_series_covid19_deaths_US.csv")

urls <- str_c(url, files)
us_cases <-read.csv(urls[1],  check.names=FALSE)
us_deaths <-read.csv(urls[2],  check.names=FALSE)
```

## Cleaning 

Next, re-orient the columns so they are more readable, and remove unnecessary attributes:

```{r usCases, echo=FALSE}
us_cases <- us_cases %>%
    pivot_longer(cols = -(UID:Combined_Key), 
                 names_to = "date", 
                 values_to = "cases") %>%
  select(Admin2:cases) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat,Long_))
```

```{r usDeaths, echo=FALSE}
us_deaths <- us_deaths %>%
    pivot_longer(cols = -(UID:Population), 
                 names_to = "date", 
                 values_to = "deaths") %>%
  select(Admin2:deaths) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat,Long_))
```

Can join the cases and deaths since most of the other data is the same:
```{r join, echo=FALSE}
us_covid <- us_deaths %>%
  left_join(us_cases, by = c("date", "Combined_Key", "Province_State")) %>%
  select(c(Combined_Key, Population,date,deaths, cases,Province_State))

```

# Analysis 

## US Daily: 
```{r us_summary, echo=FALSE}
us_overall <- us_covid %>%
  group_by(date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  mutate(cases_per_mill = cases * 1000000 / Population) %>%
  select(date, cases, deaths, deaths_per_mill, cases_per_mill, Population) %>%
  ungroup()

us_overall <-us_overall %>%
    mutate(new_cases = cases - lag(cases),  new_deaths = deaths - lag(deaths))
```

## In Colorado 
```{r colorado, echo=FALSE}

colorado_cases <- us_covid %>%
  filter(Province_State == "Colorado")

```

```{r colorado_cases, echo=FALSE}

colorado_overall <- colorado_cases %>%
  group_by(date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  mutate(cases_per_mill = cases * 1000000 / Population) %>%
  select(date, cases, deaths, deaths_per_mill, cases_per_mill, Population) %>%
  ungroup()

colorado_overall <-colorado_overall %>%
    mutate(new_cases = cases - lag(cases),  new_deaths = deaths - lag(deaths))
```

## Overall Cases US vs Colorado
```{r cases, echo=FALSE}
ggplot() + 
geom_point(data=colorado_overall, aes(x = date, y = cases_per_mill, color="Colorado")) + 
geom_point(data=us_overall, aes(x = date, y = cases_per_mill, color="US")) + 
theme(legend.position = "bottom", 
      axis.text.x = element_text(angle = 90)) + 
  xlab("Date") + 
  ylab("Cases Per Million") +
labs(title = str_c("COVID19 Cases"))
```

## Overall Deaths US vs Colorado

```{r deaths, echo=FALSE}
ggplot() + 
geom_point(data=colorado_overall, aes(x = date, y = deaths_per_mill, color="Colorado")) + 
geom_point(data=us_overall, aes(x = date, y = deaths_per_mill, color="US")) + 
  xlab("Date") + 
  ylab("Deaths Per Million") +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(title = "COVID19 Deaths")
```


## New Daily Cases US vs Colorado
```{r daily_cases, echo=FALSE}
ggplot() + 
geom_point(data=colorado_overall, aes(x = date, y = new_cases, color="Colorado")) + 
geom_point(data=us_overall, aes(x = date, y = new_cases, color="US")) + 
scale_y_log10() + 
    xlab("Date") + 
  ylab("New Cases") +
theme(legend.position = "bottom", 
      axis.text.x = element_text(angle = 90)) + 
labs(title = str_c("New Daily Cases"))
```

## New Daily Deaths US cs Colorado
```{r daily_deaths, echo=FALSE}
ggplot() + 
geom_point(data=colorado_overall, aes(x = date, y = new_deaths, color="Colorado")) + 
geom_point(data=us_overall, aes(x = date, y = new_deaths, color="US")) + 
scale_y_log10() + 
    xlab("Date") + 
  ylab("New Deaths") +
theme(legend.position = "bottom", 
      axis.text.x = element_text(angle = 90)) + 
labs(title = str_c("New Daily Deaths"))
```

## Modeling Colorado Cases

```{r model, echo=FALSE}
mod <- lm(deaths_per_mill ~ cases_per_mill, data = colorado_overall)
summary(mod)


colorado_overall %>% mutate(pred = predict(mod)) 

predictions <- colorado_overall %>% mutate(pred = predict(mod))
predictions %>% 
  ggplot() + 
  geom_point(aes(x = cases_per_mill, y = deaths_per_mill, color = "Actual")) + 
  geom_point(aes(x = cases_per_mill, y = pred, color = "prediction")) +
  xlab("Cases") + 
  ylab("Deaths")+
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(title = str_c("Model of Cases v. Deaths"))

```

# Conclusions